Finding Quality Issues in SKOS Vocabularies

### 2012-04-02 analysis.tex
fixed typos
added AC to definition, changed lc computation to use only AC, updated numbers for STW 8.10 including available mappings

The basic statistical properties of our vocabulary selection are summarized in Table~\ref{tab:vocabs}: the number of concepts, ``authoritative concepts'', as explained in Section~\ref{subsec:ld_issues}, all \texttt{skos:prefLabel}, \texttt{skos:altLabel}, and \texttt{skos:hiddenLabel} relations involving concepts (\texttt{Concept Labels}), all asserted semantic relations, as well as the number of concept schemes. From these properties we can, for instance, already see that approximately 3,000 \texttt{DBpedia Categories} concepts do not have labels (e.g., \url{Category:South_Korean_social_scientists}), which indicates that the same amount of Wikipedia categories misses natural language descriptions.

@@

\textbf{Vocabulary} & \rotatebox{90}{\textbf{Abbreviation}} & \rotatebox{90}{\textbf{\parbox{2.5cm}{Version/\\Last Modified}}} & \rotatebox{90}{\textbf{Concepts}} & \rotatebox{90}{\textbf{\parbox{2.5cm}{Authoritative\\Concepts}}} & \rotatebox{90}{\textbf{\parbox{2.5cm}{Concept\\Labels}}} & \rotatebox{90}{\textbf{\parbox{2.5cm}{Semantic\\Relations}}} & \rotatebox{90}{\textbf{\parbox{2.53cm}{Concept\\Schemes}}} \\
\toprule
United Nations Agricultural Thesaurus & \textbf{AGROVOC} & 1.3 & 32,035 & 32,035 & 620,629 & 65,934 & 1 \\
\hline
DBpedia Categories & \textbf{DBpedia} & 3.7 & 743,410 & 743,410 & 740,352 & 1,490,316 & 0 \\
\hline
The EU's Multilingual Thesaurus & \textbf{Eurovoc} & 5.0 & 6,797 & 6,797 & 457,788 & 18,491 & 128 \\
\hline
Geonames Ontology & \textbf{Geonames} & 2.2.1 & 671 & 671 & 671 & 0 & 9 \\
\hline
Gemeenschappelijke Thesaurus Audiovisuele Archieven & \textbf{GTAA} & 2010/08/25 & 171,991 & 171,991 & 178,776 & 50,892 & 9 \\
\hline
Integrated Public Sector Vocabulary & \textbf{IPSV} & 2.00 & 4,732 & 3,080 & 7,945 & 13,843 & 3 \\
\hline
Library of Congress Subject Headings & \textbf{LCSH} & 2012/03/29 & 443,164 & 408,009 & 750,219 & 598,134 & 1 \\
\hline
Austrian Armed Forces Thesaurus & \textbf{LVAk} & 0.9 & 13,411 & 13,411 & 17,250 & 16,346 & 0 \\
\hline
Middle Kingdom Tombs of Ancient Egypt Thesaurus & \textbf{Meketre} & 2011/07/07 & 422 & 422 & 569 & 1,698 & 2 \\
\hline
Medical Subject Headings & \textbf{MeSH} & \cite{van2006method} & 24,626 & 24,626 & 150,617 & 38,858 & 0 \\
\hline
North American Industry Classification System & \textbf{NAICS} & 2012 & 4,175 & 2,213 & 0 & 8,684 & 1 \\
\hline
New York Times People & \textbf{NYTP} & 2010/06/22 & 4,979 & 4,979 & 4,979 & 0 & 1 \\
\hline
University of Southampton Pressinfo & \textbf{Pressinfo} & 2011/02/24 & 1,125 & 1,125 & 0 & 0 & 0 \\
\hline
Peroxisome Knowledge Base & \textbf{PXV} & 1.6 & 2,112 & 1,686 & 3,628 & 2,695 & 1 \\
\hline
Thesaurus for Economics & \textbf{STW} & 8.06 & 6,524 & 6,524 & 31,189 & 57,907 & 1 \\
\bottomrule
\end{tabular}
}
\end{center}
\end{table}

@@

The results of this analysis are summarized in Table~\ref{tab:results}, which shows the absolute number of possibly affected resources for each quality checking function and vocabulary combination. n/a means that a certain function could not be computed with the current version of qSKOS because of algorithm complexity or missing support of SKOS-XL. Numbers marked with an asterisk (*) were obtained by extrapolating from subsets containing 5\% of the respective vocabulary resources.

@@

\toprule
Omitted or Invalid Language Tags & 0 & 0 & 219 & 0 & 0 & 0 & 100,316 & 13,411 & 0 & 23,950 & n/a & 0 & 1,224 & 1,578 & 2 \\

Incomplete Language Coverage & 32,035 & 0 & 6370 & 0 & 0 & 0 & 0 & 0 & 420 & 0 & n/a & 0 & 0 & 0 & 6,456 \\

Undocumented Concepts & 32,035 & 743,410 & 5,341 & 0 & 96,850 & 4,551 & 342,848 & 13,411 & 422 & 1,807 & 3,259 & 4,094 & 1,125 & 1,918 & 5,236 \\

Label Conflicts & 2,949 & 0 & 48 & 18 & 12,404 & 0 & 11,975 & 13 & 4 & 0 & n/a & 0 & 0 & 7 & 5 \\

@@

Orphan Concepts & 0 & 77,062 & 7 & 671 & 162,000 & 0 & 173,149 & 21 & 0 & 0 & 0 & 4,979 & 1,125 & 2 & 4 \\

Weakly Connected Components & 4 & 1,506 & 4 & 0 & 621 & 1 & 22,343 & 11 & 5 & 4 & 1 & 0 & 0 & 10 & 1 \\

@@

Valueless Associative Relations & 282 & 8,839 & 6 & 0 & 9,448 & 253 & 0 & 5 & 0 & 550 & 0 & 0 & 0 & 0 & 5,139 \\

@@

Omitted Top Concepts & 0 & 0 & 1 & 9 & 9 & 0 & 1 & 0 & 0 & 0 & 0 & 1 & 0 & 0 & 0 \\

@@

Missing In-Links & 32,035 & 733,800 & 6,796 & 19 & 171,980* & 3,080 & 408,000* & 13,411 & 422 & 24,625 & 2,213 & 20 & 1,125 & 1,686 & 6,516 \\

Missing Out-Links & 32,035 & 743,410 & 6,797 & 671 & 171,991 & 0 & 344,054 & 13,411 & 273 & 24,626 & 1 & 0 & 1,116 & 1,046 & 6,524 \\

@@

% Weakly Connected Components
Three vocabularies (\texttt{IPSV}, \texttt{NAICS}, \texttt{STW}) consist of only one ``giant component'', which is often considered as being the ideal vocabulary structure. All other vocabularies split into several clusters of semantically related concepts, each of which represents a certain subtopic in the vocabularies. \texttt{Eurovoc}, for instance, contains of 4 weakly connected components, containing 4, 5, 6 and 6775 concepts. The three smallest components contain concepts thematically related to ``Africa'', ``educational institutions'' and ``forecasts''. Weakly connected components divide the \texttt{Meketre} vocabularies into, e.g., topics, and concepts reserved for internal use or museums. \texttt{GTAA} consists of 621 weakly connected components with the largest one being a subject thesaurus containing 8413 concepts. Most of the other components contain less than 10 concepts, representing, e.g., locations, person names, and genres. \texttt{PXV} splits into 10 topic-related weakly connected components, such as ``deficiancies'', ``defects'' or ``signals''. The \texttt{LVAk} thesaurus contains 11 concept clusters, which are obviously forgotten test data.

>>>>

The basic statistical properties of our vocabulary selection are summarized in Table~\ref{tab:vocabs}: the number of concepts and authoritative concepts, all \texttt{skos:prefLabel}, \texttt{skos:altLabel}, and \texttt{skos:hiddenLabel} relations involving concepts (\texttt{Concept Labels}), all asserted semantic relations, as well as the number of concept schemes. From these properties we can, for instance, already see that approximately 3,000 \texttt{DBpedia Categories} concepts do not have labels (e.g., \url{Category:South_Korean_social_scientists}), which indicates that the same amount of Wikipedia categories misses natural language descriptions.

@@

\textbf{Vocabulary} & \rotatebox{90}{\textbf{Abbreviation}} & \rotatebox{90}{\textbf{\parbox{2.5cm}{Version/\\Last Modified}}} & \rotatebox{90}{\textbf{Concepts}} & \rotatebox{90}{\textbf{\parbox{2.5cm}{Authoritative\\Concepts}}} & \rotatebox{90}{\textbf{\parbox{2.5cm}{Concept\\Labels}}} & \rotatebox{90}{\textbf{\parbox{2.5cm}{Semantic\\Relations}}} & \rotatebox{90}{\textbf{\parbox{2.53cm}{Concept\\Schemes}}} \\
\toprule
United Nations Agricultural Thesaurus & \textbf{AGROVOC} & 1.3 & 32,035 & 32,035 & 620,629 & 65,934 & 1 \\
\hline
DBpedia Categories & \textbf{DBpedia} & 3.7 & 743,410 & 743,410 & 740,352 & 1,490,316 & 0 \\
\hline
The EU's Multilingual Thesaurus & \textbf{Eurovoc} & 5.0 & 6,797 & 6,797 & 457,788 & 18,491 & 128 \\
\hline
Geonames Ontology & \textbf{Geonames} & 2.2.1 & 671 & 671 & 671 & 0 & 9 \\
\hline
Gemeenschappelijke Thesaurus Audiovisuele Archieven & \textbf{GTAA} & 2010/08/25 & 171,991 & 171,991 & 178,776 & 50,892 & 9 \\
\hline
Integrated Public Sector Vocabulary & \textbf{IPSV} & 2.00 & 4,732 & 3,080 & 7,945 & 13,843 & 3 \\
\hline
Library of Congress Subject Headings & \textbf{LCSH} & 2012/03/29 & 443,164 & 408,009 & 750,219 & 598,134 & 1 \\
\hline
Austrian Armed Forces Thesaurus & \textbf{LVAk} & 0.9 & 13,411 & 13,411 & 17,250 & 16,346 & 0 \\
\hline
Middle Kingdom Tombs of Ancient Egypt Thesaurus & \textbf{Meketre} & 2011/07/07 & 422 & 422 & 569 & 1,698 & 2 \\
\hline
Medical Subject Headings & \textbf{MeSH} & \cite{van2006method} & 24,626 & 24,626 & 150,617 & 38,858 & 0 \\
\hline
North American Industry Classification System & \textbf{NAICS} & 2012 & 4,175 & 2,213 & 0 & 8,684 & 1 \\
\hline
New York Times People & \textbf{NYTP} & 2010/06/22 & 4,979 & 4,979 & 4,979 & 0 & 1 \\
\hline
University of Southampton Pressinfo & \textbf{Pressinfo} & 2011/02/24 & 1,125 & 1,125 & 0 & 0 & 0 \\
\hline
Peroxisome Knowledge Base & \textbf{PXV} & 1.6 & 2,112 & 1,686 & 3,628 & 2,695 & 1 \\
\hline
Thesaurus for Economics & \textbf{STW} & 8.10 & 25,107 & 6,789 & 58,441 & 91,816 & 3 \\
\bottomrule
\end{tabular}
}
\end{center}
\end{table}

@@

The results of this analysis are summarized in Table~\ref{tab:results}, which shows the absolute number of possibly affected resources for each quality checking function and vocabulary combination. Numbers marked with an asterisk (*) were obtained by extrapolating from subsets containing 5\% of the respective vocabulary resources.

@@

\toprule
Omitted or Invalid Language Tags & 0 & 0 & 219 & 0 & 0 & 0 & 100,316 & 13,411 & 0 & 23,950 & !0! & 0 & 1,224 & 1,578 & 2 \\

Incomplete Language Coverage & 32,035 & 0 & 6370 & 0 & 0 & 0 & 0 & 0 & 420 & 0 & !0! & 0 & 0 & 0 & 25,050 \\

Undocumented Concepts & 32,035 & 743,410 & 5,341 & 0 & 96,850 & 4,551 & 342,848 & 13,411 & 422 & 1,807 & 3,259 & 4,094 & 1,125 & 1,918 & 23,752 \\

Label Conflicts & 2,949 & 0 & 48 & 18 & 12,404 & 0 & 11,975 & 13 & 4 & 0 & 0 & 0 & 0 & 7 & 0 \\

@@

Orphan Concepts & 0 & 77,062 & 7 & 671 & 162,000 & 0 & 173,149 & 21 & 0 & 0 & 0 & 4,979 & 1,125 & 2 & 70 \\

Weakly Connected Components & 4 & 1,506 & 4 & 0 & 621 & 1 & 22,343 & 11 & 5 & 4 & 1 & 0 & 0 & 10 & 141 \\

@@

Valueless Associative Relations & 282 & 8,839 & 6 & 0 & 9,448 & 253 & 0 & 5 & 0 & 550 & 0 & 0 & 0 & 0 & 5,004 \\

@@

Omitted Top Concepts & 0 & 0 & 1 & 9 & 9 & 0 & 1 & 0 & 0 & 0 & 0 & 1 & 0 & 0 & 2 \\

@@

Missing In-Links & 32,035 & 733,800 & 6,796 & 19 & 171,980* & 3,080 & 408,000* & 13,411 & 422 & 24,625 & 2,213 & 20 & 1,125 & 1,686 & 6,781 \\

Missing Out-Links & 32,035 & 743,410 & 6,797 & 671 & 171,991 & 0 & 344,054 & 13,411 & 273 & 24,626 & 1 & 0 & 1,116 & 1,046 & 0 \\

@@

% Weakly Connected Components
Three vocabularies (\texttt{IPSV}, \texttt{NAICS}, \texttt{STW}) consist of only one ``giant component'', which is often considered as being the ideal vocabulary structure. All other vocabularies split into several clusters of semantically related concepts, each of which represents a certain subtopic in the vocabularies. \texttt{Eurovoc}, for instance, contains of 4 weakly connected components, containing 4, 5, 6 and 6775 concepts. The three smallest components contain concepts thematically related to ``Africa'', ``educational institutions'' and ``forecasts''. Weakly connected components divide the \texttt{Meketre} vocabulary into different topics, e.g., museums of concepts reserved for internal use. \texttt{GTAA} consists of 621 weakly connected components with the largest one being a subject thesaurus containing 8413 concepts. Most of the other components contain less than 10 concepts, representing, e.g., locations, person names, and genres. \texttt{PXV} splits into 10 topic-related weakly connected components, such as ``deficiancies'', ``defects'' or ``signals''. The \texttt{LVAk} thesaurus contains 11 concept clusters, which are obviously forgotten test data.

### 2012-04-02 criteria.tex
added AC to definition, changed lc computation to use only AC, updated numbers for STW 8.10 including available mappings

\spnewtheorem*{mydef}{Definition}{\bfseries}{\itshape}
\begin{mydef}[SKOS Vocabulary]
Let a SKOS vocabulary be a tuple of the form $V = \langle IR, C, SR, LV, CS \rangle$, with \(IR = I_{CEXT}(\texttt{rdfs:Resource}^\mathcal{I})\) being the set of \textbf{resources}, \(C \subseteq IR\) with \(C = I_{CEXT}(\texttt{skos:Concept}^\mathcal{I})\) being the set of \textbf{concepts}, \(SR = I_{EXT}(\texttt{skos:semanticRelation}^\mathcal{I})\) being the set of \textbf{semantic relations} associating concepts with one another, $LV \subseteq I_{CEXT}(\texttt{rdfs:Literal}^\mathcal{I})$ being the set of untyped \textbf{plain literals}, and \(CS = I_{CEXT}(\texttt{skos:ConceptScheme}^\mathcal{I})\) being the set of \textbf{concept schemes}. Further, we let $V$ be the fully entailed RDFS interpretation of the underlying RDF graph. We enrich $V$ by entailment of \texttt{owl:inverseOf} properties as well as instances of \texttt{owl:TransitiveProperty} and \texttt{owl:SymmetricProperty} defined by the formal OWL semantics of SKOS~\cite{SkosReference2008}.
\end{mydef}

@@

@@

% Origin and rationale
The SKOS Primer~\cite{Isaac2009} recommends that ``no two concepts have the same preferred lexical label in a given language when they belong to the same concept scheme''. This issue could affect application scenarios such as auto-completion, which proposes labels based on user input. Although these extra cases are acceptable for some thesauri, we generalize the above recommendation and search for all concept pairs with their respective \texttt{skos:prefLabel}, \texttt{skos:altLabel} or \texttt{skos:hiddenLabel} property values meeting a certain similarity threshold defined by a function $sim:LV \times LV \rightarrow [0,1]$.
% Function defintion
The default, built-in similarity function checks for case-insensitive string equality with a threshold equal to 1. Label conflicts can be found by iterating over all concept pairs $C \times C$, applying $sim$ to every possible label combination, and collecting those pairs with at least one label combination meeting or exceeding a specified similarity threshold. Since data on concept scheme membership may lack because of the Open World Assumption, we handle this issue under the Closed World Assumption.

@@

% Origin and rationale
When vocabularies are being published on the Web, SKOS concepts become linkable resources. Estimating the number of in-links and identifying the concepts without any in-links, can indicate the importance of a concept.
% Function definition
We compute this by considering only the set of ``authoritative concepts'' $AC \subseteq C$, which consists of all concepts that are identified by URIs in the vocabulary namespace. For example, \texttt{PXV} links \url{http://www.peroxisomekb.nl/v1.6/pxv/C000297} to \url{http://www.bioexpert.nl/mesh/2010/concept/M0028264} using \texttt{skos:closeMatch}. Both are \texttt{skos:Concept}s, but the latter is no authoritative concept because its URI does not belong to the \texttt{PXV} namespace. We estimate the number of in-links by iterating over all elements in $AC$ and querying the Sindice\footnote{\url{http://sindice.com/} indexes the Web of Data, which is composed of pages with semantic markup in RDF, RDFa, Microformats, or Microdata. Currently it covers approximately 230M documents with over 11 billion triples.} SPARQL endpoint for triples containing the concept's URI in the object part. Empty query results are indicators for missing in-links.

>>>>

\spnewtheorem*{mydef}{Definition}{\bfseries}{\itshape}
\begin{mydef}[SKOS Vocabulary]
Let a SKOS vocabulary be a tuple of the form $V = \langle IR, C, AC, SR, LV, CS \rangle$, with \(IR = I_{CEXT}(\texttt{rdfs:Resource}^\mathcal{I})\) being the set of \textbf{resources}, \(C \subseteq IR\) with \(C = I_{CEXT}(\texttt{skos:Concept}^\mathcal{I})\) being the set of \textbf{concepts}, \(AC \subseteq C\) being the set of authoritative concepts, i.e. all concepts that are identified by URIs in the vocabulary namespace, \(SR = I_{EXT}(\texttt{skos:semanticRelation}^\mathcal{I})\) being the set of \textbf{semantic relations} associating concepts with one another, $LV \subseteq I_{CEXT}(\texttt{rdfs:Literal}^\mathcal{I})$ being the set of untyped \textbf{plain literals}, and \(CS = I_{CEXT}(\texttt{skos:ConceptScheme}^\mathcal{I})\) being the set of \textbf{concept schemes}. Further, we let $V$ be the fully entailed RDFS interpretation of the underlying RDF graph. We enrich $V$ by entailment of \texttt{owl:inverseOf} properties as well as instances of \texttt{owl:TransitiveProperty} and \texttt{owl:SymmetricProperty} defined by the formal OWL semantics of SKOS~\cite{SkosReference2008}.
\end{mydef}

@@

% where to put the example for authoritative concepts?
%For example, \texttt{PXV} links \url{http://www.peroxisomekb.nl/v1.6/pxv/C000297} to \url{http://www.bioexpert.nl/mesh/2010/concept/M0028264} using \texttt{skos:closeMatch}. Both are \texttt{skos:Concept}s, but the latter is no authoritative concept because its URI does not belong to the \texttt{PXV} namespace.

@@

% Origin and rationale
The SKOS Primer~\cite{Isaac2009} recommends that ``no two concepts have the same preferred lexical label in a given language when they belong to the same concept scheme''. This issue could affect application scenarios such as auto-completion, which proposes labels based on user input. Although these extra cases are acceptable for some thesauri, we generalize the above recommendation and search for all concept pairs with their respective \texttt{skos:prefLabel}, \texttt{skos:altLabel} or \texttt{skos:hiddenLabel} property values meeting a certain similarity threshold defined by a function $sim:LV \times LV \rightarrow [0,1]$.
% Function defintion
The default, built-in similarity function checks for case-insensitive string equality with a threshold equal to 1. Label conflicts can be found by iterating over all concept pairs $AC \times AC$, applying $sim$ to every possible label combination, and collecting those pairs with at least one label combination meeting or exceeding a specified similarity threshold. Since data on concept scheme membership may lack because of the Open World Assumption, we handle this issue under the Closed World Assumption.

@@

% Origin and rationale
When vocabularies are being published on the Web, SKOS concepts become linkable resources. Estimating the number of in-links and identifying the concepts without any in-links, can indicate the importance of a concept.
% Function definition
We estimate the number of in-links by iterating over all elements in $AC$ and querying the Sindice\footnote{\url{http://sindice.com/} indexes the Web of Data, which is composed of pages with semantic markup in RDF, RDFa, Microformats, or Microdata. Currently it covers approximately 230M documents with over 11 billion triples.} SPARQL endpoint for triples containing the concept's URI in the object part. Empty query results are indicators for missing in-links.
